\documentclass[letterpaper,11pt]{article}

\usepackage{latexsym}
\usepackage[empty]{fullpage}
\usepackage{titlesec}
\usepackage{marvosym}
\usepackage[usenames,dvipsnames]{color}
\usepackage{verbatim}
\usepackage{enumitem}
\usepackage[pdftex]{hyperref}
\usepackage{fancyhdr}
\usepackage{fontawesome5}


\pagestyle{fancy}
\fancyhf{} % clear all header and footer fields
\fancyfoot{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

% Adjust margins
\addtolength{\oddsidemargin}{-0.375in}
\addtolength{\evensidemargin}{-0.375in}
\addtolength{\textwidth}{1in}
\addtolength{\topmargin}{-.5in}
\addtolength{\textheight}{1.0in}

\urlstyle{same}

\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

% Sections formatting
\titleformat{\section}{
  \vspace{-4pt}\scshape\raggedright\large
}{}{0em}{}[\color{black}\titlerule \vspace{-5pt}]

%-------------------------
% Custom commands
\newcommand{\resumeItem}[2]{
  \item\small{
    \textbf{#1}{: #2 \vspace{-2pt}}
  }
}

\newcommand{\resumeSubheading}[4]{
  \vspace{-1pt}\item
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \textbf{#1} & #2 \\
      \textit{\small#3} & \textit{\small #4} \\
    \end{tabular*}\vspace{-5pt}
}

\newcommand{\resumeSubItem}[2]{\resumeItem{#1}{#2}\vspace{-4pt}}

\newcommand{\resumeItemNoColon}[2]{
  \item\small{
    \textbf{#1}{ #2 \vspace{-2pt}}
  }
}

\newcommand{\resumeSubItemNoColon}[2]{\resumeItemNoColon{#1}{#2}\vspace{-4pt}}


\renewcommand{\labelitemii}{$\circ$}

\newcommand{\resumeSubHeadingListStart}{\begin{itemize}[leftmargin=*]}
\newcommand{\resumeSubHeadingListEnd}{\end{itemize}}
\newcommand{\resumeItemListStart}{\begin{itemize}}
\newcommand{\resumeItemListEnd}{\end{itemize}\vspace{-5pt}}

%-------------------------------------------
%%%%%%  CV STARTS HERE  %%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}

%----------HEADING-----------------
\begin{center}
  \textbf{\Huge Pablo Fernando Badilla Torrealba}

  \vspace{12pt}

\end{center}


\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
  \faEnvelope\ \href{mailto:pbadilla.torrealba@gmail.com}{pbadilla.torrealba@gmail.com} & \faPhone\ +56971396025 \\
  \faLinkedin\ \href{https://www.linkedin.com/in/pablo-badilla-torrealba/}{pablo-badilla-torrealba} & \faGithub\ \href{https://github.com/pbadillatorrealba}{pbadillatorrealba}
\end{tabular*}

% ─── Professional Summary ───────────────────────────────────────────────────────
\section{Professional Summary}

Computer Engineer and Master of Computer Science specialized in Data Science, Machine
Learning, and Generative AI, with over 7 years of experience leading and implementing
production systems for banking, retail, academia, and start-ups, among others.

Has led technical teams with a focus on client management, scalable system architecture,
and mentoring, promoting the creation of high-standard products, best practices, and
solid end-to-end execution.

% ─── Work Experience ────────────────────────────────────────────────────────────
\section{Work Experience}

\textbf{Iniciativa de Datos e Inteligencia Artificial (IDIA) -- Universidad de Chile} \hfill Santiago, Chile \\
\textit{Tech Lead, Technology Transfer Area} \hfill April 2026 -- Present
\begin{itemize}[leftmargin=1.5em, itemsep=1pt, topsep=3pt]
  \item \textbf{Area Building \& Definition:} Leadership in the definition and structuring
        of the technology transfer area within IDIA, an initiative that brings together the
        main academic departments of FCFM (DIE, DII, DCC, DIM) to articulate research,
        transfer, teaching, and project capabilities.
  \item \textbf{AI-based ETL -- Ministerio de las Culturas:} Development of an automated
        indicator data extraction and transformation system (AI-based ETL) using Python,
        LLMs/Generative AI, and FastAPI, enabling the Ministry of Culture to process
        cultural indicator data automatically.
\end{itemize}

\vspace{6pt}
\textbf{Banco Falabella} \hfill Santiago, Chile \\
\textit{Data Scientist} \hfill August 2024 -- April 2026
\begin{itemize}[leftmargin=1.5em, itemsep=1pt, topsep=3pt]
  \item \textbf{Products \& Benefits:} AI Agent-based chat for queries about products, benefits,
        operations, and branches in Chile, Colombia, Mexico, and Peru. 100k+ monthly interactions,
        80\% cost reduction, improved latency and user experience. Internal client relationship
        management across all countries in collaboration with the Digital Channels team.
  \item \textbf{Transaction Dispute:} Development and launch of a project focused on initiating
        a fraud dispute process for unreconciled transactions. First AI initiative published by the
        bank, opening the door to more complex projects and an AI-driven strategy. Collaboration
        with the Digital Channels team.
  \item \textbf{Concurrent Conversation Server Library:} Design and implementation of a library
        based on concurrent and distributed algorithms to coordinate conversations across different
        AI use cases. Foundation for new conversational AI projects.
  \item \textbf{Mentoring:} Mentoring on conceptual development of AI conversational agents,
        technologies (LangChain / LangGraph / LangFuse, RAG, etc.), backend technologies
        (FastAPI, SQLAlchemy, Pydantic, Asyncio, etc.), and AI-assisted development (Copilot)
        for the team.
\end{itemize}

\vspace{6pt}
\textbf{Walmart Chile} \hfill Santiago, Chile \\
\textit{Machine Learning Engineer} \hfill April 2023 -- April 2024
\begin{itemize}[leftmargin=1.5em, itemsep=1pt, topsep=3pt]
  \item \textbf{Call Center Timeseries:} Development and deployment of a Machine Learning
        timeseries model for daily demand forecasting at the Call Center. Error reduced from
        60\% to 20\%. Evaluated via A/B testing and shadow deployment with ML metrics and
        cost-related KPIs. Milestone planning and joint presentation of progress and results
        to internal operations team clients together with the PO.
  \item \textbf{Queue Wait-Time Simulation:} Optimization of a project measuring queue wait
        times by store format (Lider, Express, Acuenta, Central Mayorista) and checkout type
        (MM1 and MMS) using queuing theory. Substantial architectural and performance
        refactoring of the system.
  \item \textbf{Customer Lifetime Value (CLV):} Reimplementation of the Customer Lifetime
        Value model for the Marketing area, enabling personalized campaigns by detected
        customer segment.
\end{itemize}

\vspace{6pt}
\textbf{DashAI} \hfill Santiago, Chile \\
\textit{Tech Lead} \hfill January 2023 -- June 2024
\begin{itemize}[leftmargin=1.5em, itemsep=1pt, topsep=3pt]
  \item \textbf{Development Team Leadership:} Technical leadership of a team of 5 developers
        (Universidad de Chile thesis students). Responsible for guiding various projects and
        initiatives on technical and architectural aspects.
  \item \textbf{Management \& Coordination:} Team direction, ceremony facilitation (dailies,
        plannings, refinements, etc.), definition of technical and integration requirements.
        Repository: \href{https://github.com/DashAISoftware/DashAI}{github.com/DashAISoftware/DashAI}
  \item \textbf{Funding Application \& Technology Transfer:} FONDEF 2023 grant application.
        Business model generation and plan to productize the application and its conceptual
        framework. Technology transfer strategy design, competitor analysis, and business
        model development.
  \item \textbf{Core Component Development:} Creation of Component Registry (Plugins),
        dependency injection, subproject integration, mass testing, initial front-end, and
        development best practices (CI, tests, static code verification, typing), etc.
\end{itemize}

\vspace{6pt}
\textbf{BCI} \hfill Santiago, Chile \\
\textit{Data Scientist} \hfill February 2023 -- May 2023
\begin{itemize}[leftmargin=1.5em, itemsep=1pt, topsep=3pt]
  \item \textbf{Credit Risk Assessment Features:} Creation of features based on customer
        transactional behavior to explore improvements in credit risk assessment models.
\end{itemize}

\vspace{6pt}
\textbf{National Center for Artificial Intelligence -- CENIA} \hfill Santiago, Chile \\
\textit{Machine Learning Engineer} \hfill January 2022 -- January 2023
\begin{itemize}[leftmargin=1.5em, itemsep=1pt, topsep=3pt]
  \item \textbf{End-to-End ML Model for VRP:} Design and deployment of a Machine Learning
        model that incorporates user preferences into solutions for the Vehicle Routing Problem
        (VRP). Client communication, progress presentations, milestone and delivery management
        in collaboration with the Optimization team. Delivery of a complete package ready for
        deployment in internal systems.
\end{itemize}

\vspace{6pt}
\textbf{Universidad de Chile} \hfill Santiago, Chile \\
\textit{Part-Time Lecturer} \hfill March 2021 -- July 2024
\begin{itemize}[leftmargin=1.5em, itemsep=1pt, topsep=3pt]
  \item \textbf{Scientific Programming Laboratory for Data Science:} Lecturer for the
        Scientific Programming Laboratory for Data Science course in the Master in Data Science
        (MDS) program at Universidad de Chile.
  \item \textbf{Data Handling and Extraction in Python:} Lecturer for the Data Handling and
        Extraction in Python course in the Python Applied to Data Science Diploma at the
        Department of Computer Science, Universidad de Chile.
\end{itemize}

\vspace{6pt}
\textbf{Kwiyx SPA} \hfill Santiago, Chile \\
\textit{Full Stack Developer} \hfill April 2020 -- January 2021
\begin{itemize}[leftmargin=1.5em, itemsep=1pt, topsep=3pt]
  \item \textbf{Real Estate Investment App:} Front-end development of a real estate investment
        application.
  \item \textbf{Hospital HR Management:} Full-stack development of an application for human
        resources management in pandemic hospitals based on shift optimization.
\end{itemize}

\vspace{6pt}
\textbf{Freelance Developer} \hfill Santiago, Chile \\
\textit{Data Scientist \& Full Stack Developer} \hfill July 2019 -- January 2021
\begin{itemize}[leftmargin=1.5em, itemsep=1pt, topsep=3pt]
  \item \textbf{Topic Detection System -- Universidad de Chile \& Fundación Tribu:}
        Construction of a topic detection system using Transformer embeddings to analyze the
        2019 participatory process of ``Lxs 400'' around the new Chilean Constitution.
  \item \textbf{Automated Text Detection -- Sotta y CIA:} Development of an OCR + Regex-based
        system for automatic text extraction from automotive credit promissory notes.
\end{itemize}

\vspace{6pt}
\textbf{Acción Circular} \hfill Santiago, Chile \\
\textit{Full Stack Developer} \hfill April 2019 -- July 2019
\begin{itemize}[leftmargin=1.5em, itemsep=1pt, topsep=3pt]
  \item \textbf{Recyclable Waste Management System:} Design and implementation of software
        for the management of recyclable waste collection.
\end{itemize}

\vspace{6pt}
\textbf{Cognus Consulting} \hfill Santiago, Chile \\
\textit{Data Analyst Intern} \hfill December 2017 -- February 2018
\begin{itemize}[leftmargin=1.5em, itemsep=1pt, topsep=3pt]
  \item \textbf{Data Warehousing \& Business Intelligence Solution:} Construction of a
        Data Warehousing, Business Intelligence, and visualization solution using Pentaho.
\end{itemize}

% ─── Projects ───────────────────────────────────────────────────────────────────
\section{Projects}

\textbf{Word Embeddings Fairness Evaluation Framework} (January 2020 -- June 2025): \\
Leadership and development of a library for measuring and mitigating gender, ethnic, and
religious biases in word embedding models (precursors to language models). 2 published
papers, 182 GitHub stars. \\
Repository: \href{https://github.com/dccuchile/wefe}{github.com/dccuchile/wefe} \quad
Documentation: \href{https://wefe.readthedocs.io}{wefe.readthedocs.io}

% ─── Education ──────────────────────────────────────────────────────────────────
\section{Education}

\textbf{Master of Computer Science} \hfill Santiago, Chile \\
\textit{Universidad de Chile} \hfill 2019 -- 2020

\vspace{4pt}
\textbf{Computer Engineering (Civil)} \hfill Santiago, Chile \\
\textit{Universidad de Chile} \hfill 2013 -- 2019

% ─── Publications ───────────────────────────────────────────────────────────────
\section{Publications}

\textbf{WEFE: A Python Library for Measuring and Mitigating Bias in Word Embeddings} \\
Pablo Badilla, Felipe Bravo-Marquez, María José Zambrano, Jorge Pérez. \\
\textit{Journal of Machine Learning Research}, Vol. 26, No. 156, pp. 1--6, 2025. \\
\href{http://jmlr.org/papers/v26/22-1133.html}{http://jmlr.org/papers/v26/22-1133.html}

\vspace{6pt}
\textbf{WEFE: The Word Embeddings Fairness Evaluation Framework} \\
Pablo Badilla, Felipe Bravo-Marquez, Jorge Pérez. \\
\textit{Proceedings of the Twenty-Ninth International Joint Conference on Artificial
Intelligence (IJCAI-20)}, pp. 430--436, 2020. \\
\href{https://doi.org/10.24963/ijcai.2020/60}{https://doi.org/10.24963/ijcai.2020/60}

% ─── Skills & Tools ─────────────────────────────────────────────────────────────
\section{Skills \& Tools}

\begin{itemize}[leftmargin=1.5em, itemsep=2pt, topsep=3pt]
  \item \textbf{Languages:} Spanish (Native), English B2 TOEFL-ITP
  \item \textbf{Programming Languages:} Python (Advanced), JavaScript (Intermediate)
  \item \textbf{Generative AI:} LangChain, LangGraph, LangFuse, OpenAI API, Gemini, Azure
  \item \textbf{Data Science:} NumPy, Pandas, SQL, BigQuery, Spark, JupyterLab, Plotly, Databricks
  \item \textbf{Machine Learning:} Scikit-Learn, XGBoost, LightGBM, PyTorch, HuggingFace
  \item \textbf{Backend Development:} FastAPI, SQLAlchemy, Pydantic, Asyncio, Concurrency, Dependency Injection
  \item \textbf{Cloud \& Infrastructure:} Google Cloud Platform, Vertex AI, Docker, Kubeflow
  \item \textbf{Tools \& Methodologies:} Kedro, GitHub Actions, GitLab CI, GitHub Copilot, MLOps, Scrum
  \item \textbf{Operating Systems \& Office:} Windows, Linux, macOS, Microsoft Office, Google Workspace, SharePoint, OneDrive
\end{itemize}

% ─── Additional Information ──────────────────────────────────────────────────────
\section{Additional Information}

\begin{itemize}[leftmargin=1.5em, itemsep=2pt, topsep=3pt]
  \item Graduated from Engineering and Master's with Highest Distinction.
  \item Outstanding Student 2016 and 2018: Faculty of Physical and Mathematical Sciences,
        Universidad de Chile.
  \item Thesis Supervisor: Implementation of a bias mitigation module for word embedding
        models in WEFE.
  \item Guest Lecturer: Talks on Machine Learning Operations for various undergraduate and
        graduate courses at Universidad de Chile.
  \item \textbf{Microsoft College Code Competition:} Third place in the Microsoft College Code
        Competition 2018.
  \item \textbf{Paper Reviewer:} Secondary reviewer at The Joint Conference of the 59th Annual
        Meeting of the Association for Computational Linguistics and the 11th International Joint
        Conference on Natural Language Processing (ACL-IJCNLP 2021).
\end{itemize}

\end{document}
